---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

Similarly, deal with the returns of characters.

Based on these datasets calculate the average number of deaths an Avenger suffers. 

```{r}
# Import tidyverse and other stuff
library(tidyverse)
library(readr)

# Convert the Death[1-5] columns to long format
deaths <- av %>%
  gather(key = "Time", value = "Death", starts_with("Death")) %>%
  mutate(Time = parse_number(Time))

# Convert the Return[1-5] columns to long format (if they exist in the dataset)
returns <- av %>%
  gather(key = "Time", value = "Return", starts_with("Return")) %>%
  mutate(Time = parse_number(Time))

# Calculate the average number of deaths an Avenger suffers
avg_deaths <- deaths %>%
  group_by(Name.Alias) %>%
  summarize(total_deaths = sum(Death == "YES")) %>%
  summarize(avg_deaths = mean(total_deaths))

# Print, "Average deaths of an Avenger: {value of avg_deaths}" as a string
print(paste("Average deaths of an Avenger: ", avg_deaths))
```

<!-- Average number of deaths of an avenger: ~0.54% -->

## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.


```

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.

> Logan Jorgensen: The article states that "The MVP of the Earth-616 Marvel Universe Avengers has to be Jocasta — an android based on Janet van Dyne and built by Ultron — who has been destroyed five times and then recovered five times.

> David Szczepanik: The article states that "(I) counted 89 total deaths - some unlucky Avengers are basically Meat Loaf with an E-ZPass - and on 57 occasions the individual made a comeback."

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement
```{r}
# Logan Jorgensen's code
returns %>%
  group_by(Name.Alias) %>%
  summarize(total_returns = sum(Return == "YES")) %>%
  arrange(desc(total_returns)) %>%
  head(1)
```

```{r}
# David Szczepanik's code
  death_columns <- c("Death")
  no_count <- sum(deaths[, death_columns] == "YES")
  cat("Number of deaths: ", no_count, "\n")

```


### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Logan Jorgensen: The result of the code shows that the only Avenger to be destroyed and recovered five times is Jocasta. Therefore, the statement is true.

David Szczepanik: The code above shows that there were indeed 89 deaths within the universe. The statement is true.

Upload your changes to the repository. Discuss and refine answers as a team.
